Emittance preservation of an electron bunch in a loaded quasi-linear plasma wakefield

### 2017-08-10 Tex/Paper.tex
Some edits and additions

\begin{figure*}[hbt]
    \includegraphics[width=0.495\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamEmittance}
    \includegraphics[width=0.495\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamEmittanceOffset}
    \caption{\label{Fig:BeamEmitt} The red lines show a moving window calculation of transverse normalised emittance for
        an on-axis beam with respect to the drive beam axis (left) and an offset beam (right) with an offset of one
        $\sigma_{x} = 5.24\unit{\mu m}$ in the x plane. The moving window for emittance calculation is longitudinal
        slices of $l = 4\times\Delta\xi = 18.75\unit{\mu m}$ with a $\Delta\xi$ resolution. Only slices with more than
        $100$ macro particles have been included, which accounts for the abrupt edge as well as the noise at either end
        of the beam due to low statistics. The blue lines show the peak electron beam density profile. For reference,
        the plasma density profile is included in green, but scaled up by a factor of $100$ to be visible. The solid
        lines are at the plasma entry point, the dashed lines after $4\unit{m}$ of plasma, the dash-dotted after
        $40\unit{m}$, and the dotted after $100\unit{m}$. In order to sustain a stable accelerating field for the
        witness beam, these simulations were run with an LHC energy drive beam of $7\unit{TeV}$ to avoid dephasing which
        causes the structure to brake down after about $50\unit{m}$.}
\end{figure*}

\begin{figure*}[hbt]
    \includegraphics[width=\linewidth,trim={0mm 0mm 0mm 0mm},clip]{figures/beamFilamentationAll}
    %~ \includegraphics[height=22.2mm,trim={ 0mm 9.5mm  6mm 0mm},clip]{figures/beamFilamentation01}
    %~ \includegraphics[height=22.2mm,trim={16mm 9.5mm  6mm 0mm},clip]{figures/beamFilamentation02}
    %~ \includegraphics[height=22.2mm,trim={16mm 9.5mm  0mm 0mm},clip]{figures/beamFilamentation03}
    %~ \includegraphics[height=22.2mm,trim={10mm 9.5mm 15mm 0mm},clip]{figures/beamOffsetMean}
    %~ \\
    %~ \includegraphics[height=26.0mm,trim={ 0mm 0.1mm  6mm 2mm},clip]{figures/beamFilamentation04}
    %~ \includegraphics[height=26.0mm,trim={16mm 0.1mm  6mm 2mm},clip]{figures/beamFilamentation05}
    %~ \includegraphics[height=26.0mm,trim={16mm 0.1mm  0mm 2mm},clip]{figures/beamFilamentation06}
    %~ \includegraphics[height=26.0mm,trim={10mm 0.1mm 15mm 2mm},clip]{figures/beamOffsetRMS}
    \caption{\label{Fig:BeamFilament} The left plots show the transverse phase space of the electron beam at different
        plasma position. The Right side shows the macro particle mean position (top) and RMS spread (bottom). The blue
        particles and lines represent particles with position $1.40 < \xi < 1.42$. The red particles and lines
        represent particles with position $1.55 < \xi < 1.57$.}
\end{figure*}

@@

As illustrated in figure \ref{Fig:BeamPS}, our base case of a $100\unit{pC}$, $60\unit{\mu m}$ long witness beam indeed
shows a difference in energy transfer to the tail of the beam. Also the front of the beam has a large tail in momentum
space. The positioning of the witness beam is chosen to put the bulk of the charge as close to the peak accelerating
field as possible, as can be seen in figure \ref{Fig:BeamLoading}.

Due to the high charge density resulting from a very narrow beam, the witness beam's own wakefield reaches the full
non-linear regime with $n_{eb} \approx 35\times n_{0}$. The plasma region reaches full depletion close to the peak of
the electron beam, which in turn generates strong focusing fields, with gradients of $20\unit{kT/m}$ near the beam axis
(see figure \ref{Fig:PlasmaDenTWake}).

@@

The witness beam's self-generated bubble prevents the tailing end of the beam to see any measurable emittance growth.
For the majority of the cases we studied that maintained a stable accelerating structure, about $70-80\%$ of the beam
retained its initial emittance. Figure \ref{Fig:BeamEmitt} shows the $100\unit{pC}$ and $\sigma_{z} = 60\unit{\mu m}$
case extended to a $100\unit{m}$ plasma and the drive beam energy increased to $7\unit{TeV}$ (LHC energy) to prevent
dephasing. Dephasing for the SPS beam case starts to become significant after about $50\unit{m}$.

@@

Our configuration shows robustness to electron beam transverse offset as the bubble generated by the witness beam
wakefield follows the beam. The head of the beam does not benefit from this, and we see some defocusing in this region.
However, since this the proton wakefield creates a quasi linear condition, we still see the head of the beam
stabilising. The off-axis case has a larger initial emittance growth, but the growth levels off after the first few
metres and stays constant for the remainder of the plasma accelerator section.

@@

>>>>

\begin{figure*}[hbt]
    \includegraphics[width=0.495\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamEmittance}
    \includegraphics[width=0.495\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamEmittanceOffset}
    \caption{\label{Fig:BeamEmitt} The red lines show a moving window calculation of transverse normalised emittance for
        an on-axis beam with respect to the drive beam axis (left) and an offset beam (right) with an offset of one
        $\sigma_{x} = 5.24\unit{\mu m}$ in the x plane. The moving window for emittance calculation is longitudinal
        slices of $l = 4\times\Delta\xi = 18.75\unit{\mu m}$ with a $\Delta\xi$ resolution. Only slices with more than
        $100$ macro particles have been included, which accounts for the abrupt edge as well as the noise at either end
        of the beam due to low statistics. The blue lines show the peak electron beam density profile. For reference,
        the plasma density profile is included in green, but scaled up by a factor of $100$ to be visible. The solid
        lines are at the plasma entry point, the dashed lines after $4\unit{m}$ of plasma, the dash-dotted after
        $40\unit{m}$, and the dotted after $100\unit{m}$. In order to sustain a stable accelerating field for the
        witness beam, these simulations were run with an LHC energy drive beam of $7\unit{TeV}$ to avoid de-phasing
        which causes the structure to break down after about $50\unit{m}$.}
\end{figure*}

\begin{figure*}[hbt]
    \includegraphics[width=\linewidth,trim={0mm 0mm 0mm 0mm},clip]{figures/beamFilamentationAll}
    %~ \includegraphics[height=22.2mm,trim={ 0mm 9.5mm  6mm 0mm},clip]{figures/beamFilamentation01}
    %~ \includegraphics[height=22.2mm,trim={16mm 9.5mm  6mm 0mm},clip]{figures/beamFilamentation02}
    %~ \includegraphics[height=22.2mm,trim={16mm 9.5mm  0mm 0mm},clip]{figures/beamFilamentation03}
    %~ \includegraphics[height=22.2mm,trim={10mm 9.5mm 15mm 0mm},clip]{figures/beamOffsetMean}
    %~ \\
    %~ \includegraphics[height=26.0mm,trim={ 0mm 0.1mm  6mm 2mm},clip]{figures/beamFilamentation04}
    %~ \includegraphics[height=26.0mm,trim={16mm 0.1mm  6mm 2mm},clip]{figures/beamFilamentation05}
    %~ \includegraphics[height=26.0mm,trim={16mm 0.1mm  0mm 2mm},clip]{figures/beamFilamentation06}
    %~ \includegraphics[height=26.0mm,trim={10mm 0.1mm 15mm 2mm},clip]{figures/beamOffsetRMS}
    \caption{\label{Fig:BeamFilament} The left plots show the transverse phase space of the electron beam at different
        plasma position. The Right side shows the macro particle mean position (top) and RMS spread (bottom). The blue
        particles and lines represent particles with position $1.40\unit{\mu m} < \xi < 1.42\unit{\mu m}$. The red
        particles and lines represent particles with position $1.55\unit{\mu m} < \xi < 1.57\unit{\mu m}$.}
\end{figure*}

@@

As illustrated in figure \ref{Fig:BeamPS}, our base case of a $100\unit{pC}$, $60\unit{\mu m}$ long witness beam, as
expected, shows a difference in energy transfer to the tail of the beam compared to the centre where the $E_{z}$-field
is nearly flat. Also the front of the beam has a large tail in momentum space corresponding to a smaller acceleration
gradient. The positioning of the witness beam is chosen to put the bulk of the charge as close to the peak accelerating
field as possible, but prioritising the tail end over the front as the front is in any case in the quasi-linear regime.

Due to the high charge density resulting from a very narrow beam, the witness beam's own wakefield reaches the full
non-linear regime with $n_{eb} \approx 35\times n_{0}$. The wakefield is driven by the beam's own front. The space
charge rapidly becomes high enough to start expel plasma electrons from the beam axis, and form the characteristic
electrons sheet that defines the blowout regime. As a result the plasma region reaches full depletion close to the peak
of the electron beam, which in turn generates strong focusing fields, with gradients of $20\unit{kT/m}$ near the beam
axis (see figure \ref{Fig:PlasmaDenTWake}).

@@

The witness beam's self-generated bubble prevents the tailing end of the beam to see any significant emittance growth.
For the majority of the cases we studied that maintained a stable accelerating structure, about $70-80\%$ of the beam
retained its initial emittance. Figure \ref{Fig:BeamEmitt} shows the $100\unit{pC}$ and $\sigma_{z} = 60\unit{\mu m}$
case extended to a $100\unit{m}$ plasma and the drive beam energy increased to $7\unit{TeV}$ (LHC energy) to prevent
de-phasing. De-phasing for the SPS beam case starts to become significant after about $50\unit{m}$.

@@

Our configuration also shows some robustness to electron beam transverse offset as the bubble generated by the witness
beam wakefield follows the beam. The head of the beam does not benefit from this effect, and we see some defocusing in
this region (see bottom, right plot of figure \ref{Fig:BeamFilament}). However, since the proton wakefield creates a
quasi linear condition, we still see the head of the beam stabilising (see top, right plot of figure
\ref{Fig:BeamFilament}). The off-axis case has a larger initial emittance growth (left side plot of figure
\ref{Fig:BeamFilament}), but the growth levels off after the first few metres and stays constant for the remainder of
the plasma accelerator section (figure \ref{Fig:BeamEmitt}).

@@

For accelerator applications in general, and also for Run II of the AWAKE experiment, it is desirable to maximise the
charge that can be successfully accelerated in the witness bunch. However, since longitudinally constant accelerating
fields depend on both the position and the charge density of the witness beam \cite{katsouleas:1987, tzoufras:2009}.

The large number of parameters involved makes the problem difficult to solve in simulations, and difficult to measure in
experiments. For this study we limited ourselves to look at a range of initial emittances, which allows for a range of
transverse beam sizes when with a fixed $\beta$-function matched to the plasma density; and a range of beam lengths with
a maximum at the same point on the wakefield as for our reference case covered above.
